---
title: "DATA 608 - Final Project"
author: "Ahmed Sajjad"
date: "December 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA, echo=TRUE, message = FALSE, warnings = FALSE)
```
```{r set-options, echo=FALSE, cache=FALSE}
options(width = 3000)
```

```{r}
library(dplyr)
library(stringr)
```
### <span style="color:red"><b>Final Project</b></span>
#### <span style="color:blue"><b>1) Objective:</b></span><br>
To create a Shiny App/visualization that displays the relationship between different occupations and salary for many years (2007 to 2015).

#### <span style="color:blue"><b>2) Description:</b></span><br>
 In this project we would like to find those occupations that pay higher wages and salaries. 

#### <span style="color:blue"><b>3) Data Source:</b></span><br>
Datasets related to occupation and wage estimates are available in the public domain and it can be downloaded from the following site for the years 2007 to 2015:
<li> Department of Labor Statistics (https://www.bls.gov/oes/tables.htm) </li>

#### <span style="color:blue"><b>4) Research Questions:</b></span><br>
The following questions will be addressed by this project:
<li> What are the top 30 occupations reported (for all years)</li>
<li> For a specific year, find the job title that was paid the highest in a specific occupation</li> 
<li> Find the salary history of a job title to determine if it is increasing or decreasing</li> 

#### <span style="color:blue"><b>5) Shiny App:</b></span><br>
The shiny app consists of the following tabs.
<li> Top Occupations/All Categories</li>
<li> Top Occupations/Per Category</li>
<li> Salary History/Occupation</li>

#### <span style="color:blue"><b>6) Data Cleaup/Preparation:</b></span><br>
The following functions are used to prepare data for the shiny app. All the CSV files for years 2007 to 2015 and read and cleaned.

#### <span style="color:blue"><b>7) CSV Files:</b></span><br>
This process generates the following CSV files that are used by the Shiny app:
<li> occupation_df.csv</li>
<li> salary_major_df.csv</li>
<li> salary_total_df.csv</li>
<li> title_df.csv</li>
<li> year_df.csv</li>
<br>
```{r}
# Function to convert the first Character to Uppercase
simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2), sep="", collapse=" ")
}

# Function to clean the data
getCleanData <- function(salary_year_df) {
  salary_year_df <- select(salary_year_df, matches("OCC_CODE|TITLE|OCC_GROUP|TOTAL_EMPLOYEES|SALARY|YEAR"))
  salary_year_df$OCC_CODE = gsub("\\-", "_", salary_year_df$OCC_CODE)
  salary_year_df$TITLE = sapply(salary_year_df$TITLE, simpleCap)
  salary_year_df$TOTAL_EMPLOYEES <- as.numeric(gsub(",", "", salary_year_df$TOTAL_EMPLOYEES))
  salary_year_df$SALARY <- as.numeric(gsub(",", "", salary_year_df$SALARY))
  salary_year_df$TITLE  <- gsub("\\*", "", salary_year_df$TITLE)
  salary_year_df$TITLE  <- gsub("Computer And Information Scientists, Research", 
                                "Computer And Information Research Scientists", salary_year_df$TITLE)
  salary_year_df$OCC_CODE <- gsub("29_1121", "29_1181", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("21_1099", "21_1798", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("15_1011", "15_1111", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("15_1199", "15_1199", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("15_1021", "15_1131", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("15_1041", "15_1150", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("15_1051", "15_1121", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("47_4099", "47_4799", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("15_1061", "15_1141", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("51_5022", "51_5111", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("51_9199", "51_9399", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("29_2034", "29_2037", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("29_1111", "29_1141", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("41_9099", "41_9799", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("25_3099", "25_3999", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("11_3042", "11_3131", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("13_1073", "13_1151", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE <- gsub("39_6022", "39_7012", salary_year_df$OCC_CODE)
  salary_year_df$OCC_CODE_SHORT <- str_sub(as.character(salary_year_df$OCC_CODE), 1, 2)
  salary_year_df <- arrange(salary_year_df, desc(SALARY))
  return (salary_year_df)
}

# Function to get the top 20 rows of the dataset
get20Lines <- function(salary_year_df) {
  return (head(salary_year_df, 20))
}

# Function to print the dataset
printDataFrameInfo <- function(salary_year_df, numberofRows) {
  ncol(salary_year_df)
  nrow(salary_year_df)
  head(salary_year_df, numberofRows)
}

# Function to get all occupations (majors)
getOccupationData <- function(salary_total_df) {
  occupation_df <- filter(salary_total_df, OCC_GROUP  == "major")
  occupation_df$TITLE <- gsub(" Occupations", "", occupation_df$TITLE)
  occupation_df <- arrange(occupation_df, OCC_CODE)
  occupation_df <- distinct(occupation_df, OCC_CODE, .keep_all = TRUE)
  occupation_df <- arrange(occupation_df, TITLE)
  occupation_df <- select(occupation_df,  matches("OCC_CODE|TITLE|OCC_CODE_SHORT"))
  return (occupation_df)
}

# Function to get all occupation titles (not majors)
getTitleData <- function(salary_total_df) {
  title_df <- filter(salary_total_df, OCC_GROUP  != "major")
  title_df <- select(title_df, matches("OCC_CODE|TITLE|OCC_CODE_SHORT"))
  title_df <- distinct(title_df, OCC_CODE, .keep_all = TRUE)
  title_df <- arrange(title_df, TITLE)
  return (title_df)
}


# Function to get year dataframe
getYearData <- function(salary_total_df) {
  year_df   = distinct(salary_total_df, YEAR)
  colnames(year_df) = c("YEAR_KEY")
  year_df$YEAR_VALUE = year_df$YEAR_KEY
  colnames(year_df) = c("YEAR_KEY", "YEAR_VALUE")
  return (year_df)
}

# Step 1: Read the data from different files
salary_2007_df <- read.csv(file="DATA/CSV/INPUT/SAL_2007.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)
salary_2008_df <- read.csv(file="DATA/CSV/INPUT/SAL_2008.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)
salary_2009_df <- read.csv(file="DATA/CSV/INPUT/SAL_2009.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)
salary_2010_df <- read.csv(file="DATA/CSV/INPUT/SAL_2010.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)
salary_2011_df <- read.csv(file="DATA/CSV/INPUT/SAL_2011.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)
salary_2012_df <- read.csv(file="DATA/CSV/INPUT/SAL_2012.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)
salary_2013_df <- read.csv(file="DATA/CSV/INPUT/SAL_2013.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)
salary_2014_df <- read.csv(file="DATA/CSV/INPUT/SAL_2014.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)
salary_2015_df <- read.csv(file="DATA/CSV/INPUT/SAL_2015.csv", head=TRUE,  sep=",", stringsAsFactors = FALSE)

# Step 2: Clean all datasets
salary_2007_df <- getCleanData(salary_2007_df)
salary_2008_df <- getCleanData(salary_2008_df)
salary_2009_df <- getCleanData(salary_2009_df)
salary_2010_df <- getCleanData(salary_2010_df)
salary_2011_df <- getCleanData(salary_2011_df)
salary_2012_df <- getCleanData(salary_2012_df)
salary_2013_df <- getCleanData(salary_2013_df)
salary_2014_df <- getCleanData(salary_2014_df)
salary_2015_df <- getCleanData(salary_2015_df)

# Step 3: Check the contents of the cleaned dataframes
printDataFrameInfo(salary_2007_df, 20)
printDataFrameInfo(salary_2008_df, 20)
printDataFrameInfo(salary_2009_df, 20)
printDataFrameInfo(salary_2010_df, 20)
printDataFrameInfo(salary_2011_df, 20)
printDataFrameInfo(salary_2012_df, 20)
printDataFrameInfo(salary_2013_df, 20)
printDataFrameInfo(salary_2014_df, 20)
printDataFrameInfo(salary_2015_df, 20)

# Combine all dataframes
salary_total_df = rbind (salary_2007_df, salary_2008_df, salary_2009_df, salary_2010_df, salary_2011_df, salary_2012_df, salary_2013_df, salary_2014_df, salary_2015_df)
printDataFrameInfo(salary_total_df, 20)

# Create salary major dataframe
salary_major_df = filter(salary_total_df, OCC_GROUP == "major")
printDataFrameInfo(salary_major_df, 20)

# Create occupation dataframe
occupation_df = getOccupationData(salary_total_df)
printDataFrameInfo(occupation_df, 20)

# Create title dataframe
title_df = getTitleData(salary_total_df)
printDataFrameInfo(title_df, 20)

# Create year dataframe
year_df = getYearData(salary_total_df)
printDataFrameInfo(year_df, 20)

# Write all dataframes to CSV files
# Write short year datasets for homework #06 to CSVs
write.csv(get20Lines(salary_2007_df),'DATA/CSV/OUTPUT/salary_2007.csv', row.names=FALSE)
write.csv(get20Lines(salary_2008_df),'DATA/CSV/OUTPUT/salary_2008.csv', row.names=FALSE)
write.csv(get20Lines(salary_2009_df),'DATA/CSV/OUTPUT/salary_2009.csv', row.names=FALSE)
write.csv(get20Lines(salary_2010_df),'DATA/CSV/OUTPUT/salary_2010.csv', row.names=FALSE)
write.csv(get20Lines(salary_2011_df),'DATA/CSV/OUTPUT/salary_2011.csv', row.names=FALSE)
write.csv(get20Lines(salary_2012_df),'DATA/CSV/OUTPUT/salary_2012.csv', row.names=FALSE)
write.csv(get20Lines(salary_2013_df),'DATA/CSV/OUTPUT/salary_2013.csv', row.names=FALSE)
write.csv(get20Lines(salary_2014_df),'DATA/CSV/OUTPUT/salary_2014.csv', row.names=FALSE)
write.csv(get20Lines(salary_2015_df),'DATA/CSV/OUTPUT/salary_2015.csv', row.names=FALSE)

# Write salary_total_df dataframe to CSV
write.csv(salary_total_df,'DATA/CSV/OUTPUT/salary_total_df.csv', row.names=FALSE)

# Write salary_major_df dataframe to CSV
write.csv(salary_major_df,'DATA/CSV/OUTPUT/salary_major_df.csv', row.names=FALSE)

# Write occupation_df dataframe to CSV
write.csv(occupation_df,'DATA/CSV/OUTPUT/occupation_df.csv', row.names=FALSE)

# Write title_df dataframe to CSV
write.csv(title_df,'DATA/CSV/OUTPUT/title_df.csv', row.names=FALSE)

# Write year_df dataframe to CSV
write.csv(year_df,'DATA/CSV/OUTPUT/year_df.csv', row.names=FALSE)
```
